Rutgers Autonomous Aircraft Team\\Technical Report\\2010 AUVSI UAS Competition

### 2010-05-25 paper/paper.tex
all aboard

In order to mitigate a disaster such as we had last year, we decided to build another airframe which could replace the \emph{Daedalus} at short notice. We selected an Aero-Design 12 Foot Telemaster \cite{aerodesign} because it was available as a balsa kit and similar in size to the \emph{Daedalus}. The 12 Foot Telemaster has a 12ft 1in (3.68m) wingspan, and is 7ft 10in (2.38m) in length.We named it the \emph{Knight Two} after the Rutgers University mascot, the Scarlet Knight.

@@

\subsubsection{Single Board Computer}
We selected the Beagleboard \cite{beagleboard}, a single board computer made by Texas Instruments. It features a 500MHz ARM Cortex A8 processor and 256MB ram. We installed Angstrom Linux \cite{angstromlinux} on an 8GB SD card. The Beagleboard has a few logic level serial, parallel, and general purpose IO interfaces on board, but for simplicity we chose to attach all of the devices through the USB bus. The Beagleboard provides one USB host port for this purpose, but the power supply can only sustain a low current draw. To provide more current to the attached devices, we use a Belkin 7-port powered USB hub \cite{belkin7usbhub}. With a proper power supply, the USB hub will provide up to 500mA to each port on the 5v/GND circuit.

@@

Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer \cite{lsm303dlh}. Both the ITG-3200 and LSM303DLH have an all digital interface, and communicate with a microcontroller on a single I2C \cite{i2c} serial bus. The microcontroller performs the self-test on the LSM303DLH at start-up, and proceeds to send uncalibrated samples to the flight computer over its UART connection. Samples of all 9 axes (three each of gyro, accelerometer, and magnetometer) are taken and sent to the flight computer at 40Hz.

@@

The multithread approach works well for resources which are only needed by the autopilot process, but both the attitude and GPS information are needed by both the autopilot and imaging system processes. To remedy this, we use \texttt{gpsd} \cite{gpsd}, the standard Linux GPS service, to interface with the uBlox GPS device over a serial port. \texttt{Libgps}, a library for interfacing with \texttt{gpsd}, provides updates to client threads in both the autopilot and imaging processes. We handle the AHRS updates in the same way, except instead of the specialized \texttt{libgps} library we use the generic \texttt{mqueue} \cite{mqueue} service, a message passing facility provided by the Linux kernel. Each client thread opens a named queue, and the AHRS process delivers updated measurements to each queue.

@@

The physical AHRS board sends 9 uncalibrated measurements to the flight computer, but the aircraft control loop is requires in attitude and heading expressed as 3 Euler angles \cite{euler_angles} and their derivatives with respect to time. It is straightforward to estimate Euler angle derivatives from rate gyro measurements, but estimating three degrees of angular orientation from six magnetometer and accelerometer measurements is a difficult nonlinear operation. Many algorithms exist

@@

\begin{wrapfigure}{r}{0.4\textwidth}
	\vspace{-20pt}
	\centering
	\includegraphics[width=0.37\textwidth]{../images/paparazzi_gcs.jpg}
	\caption{Paparazzi GCS. Image Credit: Paparazzi Wiki \cite{paparazziweb }}
	\label{fig:paparazzi_gcs}
	\vspace{-60pt}
\end{wrapfigure}
Because we based our autopilot software on the Paparazzi project, we were able to make use of their excellent ground control software (GCS) with essentially no modifications. The Paparazzi GCS provides a map interface that allows the operator and competition judges to inspect the aircraft's position to ensure it is within the allowed airspace. It also provides attitude and altitude telemetry, and notification of lost radio contact, so the operator can alert the safety pilot to unsafe operation.

@@

\subsection{Safety Features}
Our system is designed to gracefully handle any failure while continuing to ensure the safety of those nearby. 
\begin{itemize}
	\setlength{\itemsep}{0cm}
	\setlength{\parskip}{0cm}
	\item The channel 5 toggle switch determines whether the radio control transmitter, or the autopilot, has control over the flight control servos. The radio control safety pilot may disengage the autopilot at any time by flicking this switch.
	\item If the plane loses telemetry with the ground station for more than 10 seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the ground station in hopes of reestablishing a connection.  
	\item If the connection telemetry remains dead for 30 seconds, the autopilot will bring the servos to a predefined default (throttle closed, elevator up, rudder right, ailerons left). The servos will remain in that position until the safety pilot takes control of the aircraft.
	\item If the flight computer does not provide the servo controller with a new servo output value for longer than 2 seconds, the servo controller will bring the servos to a predefined default (throttle closed, elevator up, rudder right, ailerons left).  The servos will remain in that position until the safety pilot takes control of the aircraft.
	\item During engine start, the centerman (see \ref{appendix_preflight}) may kill the engine at any time with the ignition switch located on the right side of the aircraft, behind the engine mount.
	\item Radio control channel 6 functions as a remote kill switch on the \emph{Knight Two}, disengaging the electronic ignition system of the engine. (The \emph{Daedalus} has a magneto ignition, so we only have the manual kill switch behind the engine mount)
\end{itemize}

@@

\newpage

>>>>

In order to mitigate a disaster such as we had last year, we decided to build another airframe which could replace the \emph{Daedalus} at short notice. We selected an Aero Craft Ltd. 12 Foot Telemaster \cite{aero_craft_ltd} because it was available as a balsa kit and similar in size to the \emph{Daedalus}. The 12 Foot Telemaster has a 12ft 1in (3.68m) wingspan, and is 7ft 10in (2.38m) in length.We named it the \emph{Knight Two} after the Rutgers University mascot, the Scarlet Knight.

@@

\subsubsection{Single Board Computer}
We selected the Beagleboard \cite{beagleboard}, a single board computer made by Texas Instruments. It features a 500MHz ARM Cortex A8 processor and 256MB ram. We installed Angstrom Linux \cite{angstromlinux} on an 8GB SD card. The Beagleboard has a few logic level serial, parallel, and general purpose IO interfaces on board, but for simplicity we chose to attach all of the devices through the USB bus. The Beagleboard provides one USB host port for this purpose, but the power supply can only sustain a low current draw. To provide more current to the attached devices, we use a Belkin 7-port powered USB hub. With a proper power supply, the USB hub will provide up to 500mA to each port on the 5v/GND circuit.

@@

Thanks to sensor donations from Invensense and ST Micro, we were able to build a 9 axis Attitude and Heading Reference Sensor (AHRS) from an Invensense ITG-3200 3-axis rate gyro \cite{itg3200} and an ST Micro LSM303DLH 3-axis accelerometer and magnetometer \cite{lsm303dlh}. Both the ITG-3200 and LSM303DLH have an all digital interface, and communicate with a microcontroller on a single I2C \cite{wikii2c} serial bus. The microcontroller performs the self-test on the LSM303DLH at start-up, and proceeds to send uncalibrated samples to the flight computer over its UART connection. Samples of all 9 axes (three each of gyro, accelerometer, and magnetometer) are taken and sent to the flight computer at 40Hz.

@@

The multithread approach works well for resources which are only needed by the autopilot process, but both the attitude and GPS information are needed by both the autopilot and imaging system processes. To remedy this, we use \texttt{gpsd} \cite{gpsd}, the standard Linux GPS service, to interface with the uBlox GPS device over a serial port. \texttt{libgps}, a library for interfacing with \texttt{gpsd}, provides updates to client threads in both the autopilot and imaging processes. We handle the AHRS updates in the same way, except instead of the specialized \texttt{libgps} library we use the generic \texttt{mqueue} \cite{mqueue} service, a message passing facility provided by the Linux kernel. Each client thread opens a named queue, and the AHRS process delivers updated measurements to each queue.

@@

The physical AHRS board sends 9 uncalibrated measurements to the flight computer, but the aircraft control loop is requires in attitude and heading expressed as 3 Euler angles and their derivatives with respect to time. It is straightforward to estimate Euler angle derivatives from rate gyro measurements, but estimating three degrees of angular orientation from six magnetometer and accelerometer measurements is a difficult nonlinear operation.

While their discussion is outside the scope of this paper, many algorithms exist for this conversion. We use a Quaternion Kalman Filter algorithm described in \cite{choukroun04}, implemented as a complete filter in Haskell. Angular orientation is tracked in quaternion form, and translated into euler angles for the use of the autopilot and imaging processes. This implementation was carried out as part of an independent study by team member Pat Hickey, and more details can be found in \cite{hickey10hsqkf}.

@@

\begin{wrapfigure}{r}{0.4\textwidth}
	\vspace{-20pt}
	\centering
	\includegraphics[width=0.37\textwidth]{../images/paparazzi_gcs.jpg}
	\caption{Paparazzi GCS. Image Credit: Paparazzi Wiki \cite{paparazziweb }}
	\label{fig:paparazzi_gcs}
	\vspace{-40pt}
\end{wrapfigure}
Because we based our autopilot software on the Paparazzi project, we were able to make use of their excellent ground control software (GCS) with essentially no modifications. The Paparazzi GCS provides a map interface that allows the operator and competition judges to inspect the aircraft's position to ensure it is within the allowed airspace. It also provides attitude and altitude telemetry, and notification of lost radio contact, so the operator can alert the safety pilot to unsafe operation.

@@

\subsection{Safety Features}
Our system is designed to gracefully handle any failure while continuing to ensure the safety of those nearby. 
\begin{itemize}
	\setlength{\itemsep}{0cm}
	\setlength{\parskip}{0cm}
	\item The channel 5 toggle switch determines whether the radio control transmitter, or the autopilot, has control over the flight control servos. The radio control safety pilot may disengage the autopilot at any time by flicking this switch.
	\item If the plane loses telemetry with the ground station for more than 10 seconds, the autopilot will default to a ``return home mode'' that will direct the plane towards the ground station in hopes of reestablishing a connection.  
	\item If the connection telemetry remains dead for 30 seconds, the autopilot will bring the servos to a predefined default (throttle closed, elevator up, rudder right, ailerons left). The servos will remain in that position until the safety pilot takes control of the aircraft.
	\item If the flight computer does not provide the servo controller with a new servo output value for longer than 2 seconds, the servo controller will bring the servos to a predefined default (throttle closed, elevator up, rudder right, ailerons left).  The servos will remain in that position until the safety pilot takes control of the aircraft.
	\item During engine start, the centerman (see \ref{sec:appendix_preflight}) may kill the engine at any time with the ignition switch located on the right side of the aircraft, behind the engine mount.
	\item Radio control channel 6 functions as a remote kill switch on the \emph{Knight Two}, disengaging the electronic ignition system of the engine. (The \emph{Daedalus} has a magneto ignition, so we only have the manual kill switch behind the engine mount)
\end{itemize}

@@
